Revisiting the Relationship between Short-Chain Fatty Acids, the Microbiota, and Colorectal Tumors

### 2018-08-07 submission/manuscript.Rmd
last paragraph ada comments discussion

The potential protection against colorectal cancers may not be from SCFAs, even though taxa associated with their production are higher in individuals without tumors versus patients with carcinomas [@Shah2017; @normalization_Sze2017; @meta_analysis_crc_Sze2018]. Indeed our data would support the contention that the taxa are similar to those associated with SCFA production but that these specific microbes or OTUs themselves are not associated with SCFAs. In particular, our results showing that different OTUs from the same taxonomic classification are in tumor and SCFA Random Forest models supports this hypothesis. This leads to the possibility that protection may be through two other routes. First, there could be a different pathway or other less extensively studied metabolites that provides the necessary protection against tumorigenesis. Alternatively, protection may not occur via a metabolite but instead through niche exclusion of mouth-associated microbes (e.g. *Fusobacterium*, *Porphyromonas*, *Parvimonas*, *Peptostreptococcus* [@Zeller2014; @Baxter2016; @Kostic2011]). The idea of niche exclusion is similar to how the community protects against *Clostridium difficile* infection [@Theriot2015] with chronic inflammation replacing the role of antibiotics. Even though we did not find lower concentrations of SCFAs associated with colorectal tumors, we think that there are many exciting new avenues to explore because of these results.

>>>>

The potential protection against colorectal cancers may not be from SCFAs, even though taxa associated with their production are higher in individuals without tumors versus patients with carcinomas [@Shah2017; @normalization_Sze2017; @meta_analysis_crc_Sze2018]. Our data is consistent with the literature despite finding no difference in SCFA concentrations between individuals without tumors and patients with adenomas or carcinomas. In particular, we observed that the important OTUs in our tumor models were to taxa that are associated with SCFA production but that these OTUs were not themselves associated with acetate, butyrate, or propionate production. One of the clearest results to highlight this within our study were the results showing that different OTUs from the same taxonomic classification are in tumor and SCFA Random Forest models [Figure 4 & 5]. This leads to the possibility that protection may be through other routes. First, there could be a different pathway or other less extensively studied metabolites that provide the necessary protection against tumorigenesis. Alternatively, previous research has shown that non-resident oral microbes, such as *Fusobacterium*, *Porphyromonas*, *Parvimonas*, and *Peptostreptococcus* are associated with carcinomas [@Zeller2014; @Baxter2016; @Kostic2011] and can increase inflammation [@Flynn2016]. When these non-resident oral microbes increase inflammation they change the existing environment to one in which is more hospitable for themselves [@Flynn2016]. It is possible that these resident microbes provide protection not through a metabolite but instead through active exclusion of these non-resident oral microbes which in turn prevents them from establishing themselves within the community. Overall, even though we did not find lower concentrations of SCFAs associated with colorectal tumors, we think that our results open up many new and exciting avenues for investigation into how metabolites and the microbiota can prevent tumorigenesis.

### 2018-08-07 submission/manuscript.tex
last paragraph ada comments discussion

The potential protection against colorectal cancers may not be from
SCFAs, even though taxa associated with their production are higher in
individuals without tumors versus patients with carcinomas (10, 11, 27).
Indeed our data would support the contention that the taxa are similar
to those associated with SCFA production but that these specific
microbes or OTUs themselves are not associated with SCFAs. In
particular, our results showing that different OTUs from the same
taxonomic classification are in tumor and SCFA Random Forest models
supports this hypothesis. This leads to the possibility that protection
may be through two other routes. First, there could be a different
pathway or other less extensively studied metabolites that provides the
necessary protection against tumorigenesis. Alternatively, protection
may not occur via a metabolite but instead through niche exclusion of
mouth-associated microbes (e.g. \emph{Fusobacterium},
\emph{Porphyromonas}, \emph{Parvimonas}, \emph{Peptostreptococcus} (6,
12, 13)). The idea of niche exclusion is similar to how the community
protects against \emph{Clostridium difficile} infection (28) with
chronic inflammation replacing the role of antibiotics. Even though we
did not find lower concentrations of SCFAs associated with colorectal
tumors, we think that there are many exciting new avenues to explore
because of these results.

@@

\hypertarget{ref-Theriot2015}{}
28. \textbf{Theriot CM}, \textbf{Young VB}. 2015. Interactions between
the gastrointestinal microbiome and clostridium difficile. Annual Review
of Microbiology \textbf{69}:445--461.
doi:\href{https://doi.org/10.1146/annurev-micro-091014-104115}{10.1146/annurev-micro-091014-104115}.

>>>>

The potential protection against colorectal cancers may not be from
SCFAs, even though taxa associated with their production are higher in
individuals without tumors versus patients with carcinomas (10, 11, 27).
Our data is consistent with the literature despite finding no difference
in SCFA concentrations between individuals without tumors and patients
with adenomas or carcinomas. In particular, we observed that the
important OTUs in our tumor models were to taxa that are associated with
SCFA production but that these OTUs were not themselves associated with
acetate, butyrate, or propionate production. One of the clearest results
to highlight this within our study were the results showing that
different OTUs from the same taxonomic classification are in tumor and
SCFA Random Forest models {[}Figure 4 \& 5{]}. This leads to the
possibility that protection may be through other routes. First, there
could be a different pathway or other less extensively studied
metabolites that provide the necessary protection against tumorigenesis.
Alternatively, previous research has shown that non-resident oral
microbes, such as \emph{Fusobacterium}, \emph{Porphyromonas},
\emph{Parvimonas}, and \emph{Peptostreptococcus} are associated with
carcinomas (6, 12, 13) and can increase inflammation (28). When these
non-resident oral microbes increase inflammation they change the
existing environment to one in which is more hospitable for themselves
(28). It is possible that these resident microbes provide protection not
through a metabolite but instead through active exclusion of these
non-resident oral microbes which in turn prevents them from establishing
themselves within the community. Overall, even though we did not find
lower concentrations of SCFAs associated with colorectal tumors, we
think that our results open up many new and exciting avenues for
investigation into how metabolites and the microbiota can prevent
tumorigenesis.

@@

\hypertarget{ref-Flynn2016}{}
28. \textbf{Flynn KJ}, \textbf{Baxter NT}, \textbf{Schloss PD}. 2016.
Metabolic and community synergy of oral bacteria in colorectal cancer.
mSphere \textbf{1}.
doi:\href{https://doi.org/10.1128/msphere.00102-16}{10.1128/msphere.00102-16}.
